---
title: "**forecast** Package Crash Course"
author: "Mario Annau"
date: "November 21, 2016"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(forecast)
amzn.rev <- ts(read.csv("forecast_intro/amzn_rev.csv")[, 2], start = c(2000, 1), freq = 4)
```

<!-- ## Contents -->

<!-- - Intro, Influenza (Google Trends) -->
<!-- - ~~Company Data~~ (Source?) -->
<!-- - Bike Traffic in Vienna -->
<!-- - ~~Stock Prices~~, Volatility (GARCH) -->
<!-- - ~~Birdlife Data~~ -->

## Influenza (Google Trends)

```{r influenza-1, echo = TRUE}
# from https://www.google.com/trends/explore?cat=45&date=all&geo=AT&q=influenza
dat.raw <- read.csv("forecast_intro/Influenza.csv", header = TRUE, skip = 1, stringsAsFactors = FALSE)
dat.ts <- ts(dat.raw[, 2], frequency = 12, start = c(2004, 1))
plot(dat.ts)
dat.ts
```

The time series can be decomposed into three components: seasonal, trend-cycle and a residual component.

$$y_t = S_t + T_t + E_t$$

If seasonal component depends on absolute level of time series we can also use a multiplicative model
$$y_t = S_t \times T_t \times E_t$$
By taking the logarithm the additive model can be used.

```{r influenza-2, echo = TRUE}
dat.ts.decomp <- decompose(dat.ts, type = "additive")
plot(dat.ts.decomp)
```

```{r influenza-3, echo = TRUE}
dat.ts.noseason <- dat.ts - dat.ts.decomp$seasonal
plot(dat.ts.noseason)
lines(dat.ts.decomp$trend, col = "red")
```

$$ARIMA(p,d,q)$$ models can be used and are specified as $$\dots$$
(p, d, q) are the AR order, the degree of differencing, and the MA order.

First we can check if timeseries is stationary. If not clear on plot or unsure, we can use the Augmented Dickey–Fuller Test or Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test

```{r influenza-df, echo = TRUE}
tseries::adf.test(dat.ts, alternative = "stationary")
tseries::kpss.test(dat.ts)
```

Since no differencing is required, we set $$d=0$$ and check the acf and pacf plots
```{r influenza-4, echo = TRUE}
acf(dat.ts, lag.max = 25)
pacf(dat.ts, lag.max = 25)
```

The acf plots suggests a second order autoregressive process, we thus set $$p=2$$. Let's check the current model now:
```{r influenza-5, echo = TRUE}
pacf(arima(dat.ts, order= c(2, 0, 0))$residuals)
```

Ups - the pacf plot shows that there is still a seasonal pattern left. We try it with a seasonal order AR of $$p=1$$
```{r influenza-6, echo = TRUE}
pacf(arima(dat.ts, order= c(2, 0, 0), seasonal = c(1, 0, 0))$residuals)
```

And increase to two since we still have a seasonal effect
```{r influenza-7, echo = TRUE}
pacf(arima(dat.ts, order= c(2, 0, 0), seasonal = c(2, 0, 0))$residuals)
```

This is the same model as selected by auto.arima 
```{r influenza-8, echo = TRUE}
mod1 <- auto.arima(dat.ts)
mod1
fc <- forecast(mod1, 24)
plot(fc)
fc
```

## Company Data (AMZN)

```{r company-1, echo = TRUE, eval=FALSE}
urls <- paste("http://www.wikinvest.com/stock/Amazon.com_(AMZN)/Data/Revenue", rep(2000:2015, each=4), sprintf("Q%d", 1:4), sep = "/")
vals <- sapply(urls, function(x) {
  tree <- htmlTreeParse(x, useInternalNodes = TRUE)
  xmlValue(getNodeSet(tree, "//div[@id='nv_value']")[[1]])
})
matchvals <- gregexpr("[0-9.]+", vals)
numvals <- as.numeric(sapply(1:length(vals), function(i) substr(vals[i], matchvals[[i]][1], attr(matchvals[[i]], "match.length")[1] + 1)))
numvals <- ifelse(grepl("billion", vals), numvals * 1000, numvals)
amzn.rev <- ts(numvals, start = c(2000, 1), frequency = 4)
```

```{r company-2, echo = TRUE}
plot(amzn.rev)
```

We can easily see that should 
1. Use the log scale or
2. Differentiate the series.
```{r company-3, echo = TRUE}
amzn.rev.log <- log(amzn.rev)
plot(amzn.rev.log)
```

Now it's easy to see the trend and the seasonal component. We are now lazy and use auto.arima again:
```{r company-4, echo = TRUE}
mod2 <- auto.arima(amzn.rev.log)
mod2
plot(forecast(mod2, 12))
mod3 <- auto.arima(amzn.rev)
mod3
fc3 <- forecast(mod3, 12)
plot(fc3)
fc3
```

## Reading Materials

- [Forecasting Principles and Practice](https://www.otexts.org/fpp) by Rob Hyndman and  George Athanasopoulos

