---
title: "README"
author: "ViennaR"
date: "6/8/2019"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Prerequisites

- Current version of RStudio
- Current version of R
- Git, Github account
- R-Packages: blogdown, rmarkdown, ...
- Hugo 0.55.6 (can also be installed interactively within R Console later)

```{r, eval = FALSE}
blogdown::install_hugo()
```

## Create new Project

In Rstudio, create new project

1. File -> New Project -> Version Control
1. Use the following URL: https://github.com/ViennaR/ViennaR-Blog.git
1. Create new Project

## Folder Structure

- content/post: Posts of the blog
- static: Images, slides, etc
- themes: Theme of the blog, should not be touched

## Serve Sites

In RStudio: Addins -> Serve Site

Or in the console:
```{r, eval = FALSE}
blogdown:::serve_site()
```

## Add new post

In RStudio: Run Addins -> New Post

## Publish

To publish site to viennar.org use the following steps:

1. Turn off hugo build server, if running using `servr::daemon_stop()`
1. In case README.Rmd has changed: knit README.Rmd
1. Run `blogdown::hugo_build()`
1. Commit all changes to master
1. Run `git push origin master` or simply push **Push** button




